\vspace{-0.25em}\thetitle

### 2009-07-23 paper.tex
...

To see the linear structure of these representation matrices, we apply one of the most powerful tools of linear algebra:
the singular value decomposition (\caps{SVD}).
This matrix decomposition can be applied to any real matrix, factoring it into a product of three real matrices:
\begin{align}
  X = US\trans{V}.
\end{align}
Here $U$ and $V$ are orthogonal and $S$ is diagonal, with entries of nonincreasing nonnegative value.
Among the useful properties of this factorization, is that the first $k$ columns of $US$ are an optimal $k$-dimensional reduced model of $X$.\,\svdnote
This property allows us to visualize the structure of $X$ using initial columns of $US$.
However, we are interested only interested in direction since
vectors differing only by magnitude lie in the same one-dimensional subspace.
To visualize direction while eliminating distance, we project all data points onto the unit-sum hyperplane through the origin.\projectionnote

\Figure{svd} illustrates the results of applying \caps{SVD} dimension reduction to sample matrices of six different traffic traces.
The figures show scatter plots of the most significant two projected \caps{SVD} coordinates for a matrix of 5000 sample flows from each trace.
% The original data encodes packet size distributions, inter-packet interval distributions, \caps{IP} protocol numbers, and port numbers for a sample of 5000 flows from each trace.
% The exact details of how the linear representation encodes these features will be discussed in \Section{methodology}.
Linear structure is clearly visible in every one:
flow behaviors are heavily concentrated near low-dimensional subspaces.
Points concentrated along a one-dimensional subspace appear as dark ``spots'' where many points have been plotted;
points concentrated along a two-dimensional subspace appear as lines;
points falling near subspaces of more dimensions appear as ``smears.''
% There is a smattering of outliers, not especially near such structures.

>>>>

To see the linear structure of these representation matrices, we apply one of the most powerful tools of linear algebra:
the singular value decomposition (\caps{SVD}).
This matrix decomposition can be applied to any real matrix, factoring it into a product of three real matrices:
\begin{align}
  X = US\trans{V}.
\end{align}
$U$ and $V$ are orthogonal and $S$ is diagonal, with entries of nonincreasing nonnegative value.
Among the useful properties of this factorization, is that the first $k$ columns of $US$ are an optimal $k$-dimensional reduced model of $X$.\,\svdnote
This property allows us to visualize the structure of $X$ using initial columns of $US$.
However, we are interested only interested in direction since
vectors differing only by magnitude lie in the same one-dimensional subspace.
To visualize direction while eliminating distance, we project all data points onto the unit-sum hyperplane through the origin.\projectionnote

\Figure{svd} illustrates the results of applying \caps{SVD} dimension reduction to sample matrices of six different traffic traces.
The figures show scatter plots of the most significant two projected \caps{SVD} coordinates for a matrix of 5000 sample flows from each trace.
% The original data encodes packet size distributions, inter-packet interval distributions, \caps{IP} protocol numbers, and port numbers for a sample of 5000 flows from each trace.
% The exact details of how the linear representation encodes these features will be discussed in \Section{methodology}.
Linear structure is visible in all six:
flow behaviors are heavily concentrated near low-dimensional subspaces.
Points concentrated along a one-dimensional subspace appear as dark ``spots'' where many points have been plotted;
points concentrated near two-dimensional subspaces appear as lines;
points falling along higher subspaces appear as ``smears.''
% There is a smattering of outliers, not especially near such structures.
